---
title: "Data 607 Week 10 Assignment"
author: "Leticia Salazar"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
$~$

### **Sentiment Analysis**

$~$

#### In [*Text Mining with R*](https://www.tidytextmining.com/sentiment.html), Chapter 2 looks at Sentiment Analysis. In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document. You should provide a citation to this base code. You're then asked to entend the code in two ways:

* Work with a different corpus of your choosing, and 
* Incorporate at least one additional sentiment lexicon (possibly from another R package that you've found through research).

$~$

### **From the Textbook:**

$~$

#### Below is the code example from Chapter 2 of Text Mining with R

```{r}
# load libraries
library("tidyverse")
library("tidytext")
library("janeaustenr")
library("tidyr")
library("dplyr")
library("stringr")
library("ggplot2")
```


$~$

#### **Get `sentiments` datasets**

```{r}
# get sentiments for afinn, bing and nrc

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

$~$

#### **Sentiment analysis with inner join**

```{r}
# get data tidied
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
      ignore_case = TRUE
    )))
  ) %>%
  ungroup() %>%
  unnest_tokens(word, text)


# sentiment analysis
nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```




```{r}
# Positive and Negative sentiments 

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```


```{r}
# Plot

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

$~$

#### **Comparing the three sentiment dictionaries**

```{r}
# Filter book Pride and Prejudice

pride_prejudice <- tidy_books %>%
  filter(book == "Pride & Prejudice")

pride_prejudice
```


```{r}
afinn <- pride_prejudice %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = linenumber %/% 80) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")
```



```{r}
bing_and_nrc <- bind_rows(
  pride_prejudice %>%
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>%
    inner_join(get_sentiments("nrc") %>%
      filter(sentiment %in% c(
        "positive",
        "negative"
      ))) %>%
    mutate(method = "NRC")
) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```



```{r}
bind_rows(
  afinn,
  bing_and_nrc
) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```


```{r}
get_sentiments("nrc") %>%
  filter(sentiment %in% c(
    "positive",
    "negative"
  )) %>%
  count(sentiment)
```


```{r}
get_sentiments("bing") %>%
  count(sentiment)
```


$~$

#### **Most common positive and negative words**

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```



```{r}
# Plot word counts

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(
    y = "Contribution to sentiment",
    x = NULL
  ) +
  coord_flip()

```



```{r}
# Stop words

custom_stop_words <- bind_rows(tibble(word = c("miss"), 
                                          lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```


$~$

#### **Wordclouds**

```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

```



```{r}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("gray20", "gray80"),
    max.words = 100
  )

```

$~$

#### **Looking at units beyond just words**

```{r}
PandP_sentences <- tibble(text = prideprejudice) %>%
  unnest_tokens(sentence, text, token = "sentences")

PandP_sentences$sentence[2]
```



```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text,
    token = "regex",
    pattern = "Chapter|CHAPTER [\\dIVXLC]"
  ) %>%
  ungroup()

austen_chapters %>%
  group_by(book) %>%
  summarise(chapters = n())
```



```{r}
bingnegative <- get_sentiments("bing") %>%
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())
```



```{r}
tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords / words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup()
```
$~$

### **Code Extension:**

#### I found this additional corpus called `loughran` and found it on [rdocumentation.org](https://www.rdocumentation.org/packages/tidytext/versions/0.2.3/topics/get_sentiments)

$~$

#### **[YH.C. Andersens Fairy Tales](https://github.com/EmilHvitfeldt/hcandersenr)**

$~$

```{r}
library(devtools)
install_github("EmilHvitfeldt/hcandersenr")
library(hcandersenr)

# install.packages("hcandersenr")
library("hcandersenr")
```


```{r}
#  Sentiment loughran

get_sentiments("loughran")
```

$~$

#### **Available fairy tales: first 15 books**

```{r}
# Code was taken from https://github.com/EmilHvitfeldt/hcandersenr

hca_fairytales() %>% 
  select(book, language) %>% 
  unique() %>% 
  mutate(langauge = fct_relevel(language, c("English", "Spanish", "German", "Danish", "French"))) %>%
  top_n(60, book) %>% 
  ggplot(aes(langauge, book)) + 
  geom_raster(alpha = 0.3) +
  scale_x_discrete(position = "top")
```



```{r}
tidy_books_hca <- hca_fairytales() %>%
  group_by(book) %>%
 
  ungroup() %>%
  unnest_tokens(word, text)
```



```{r}
# Filter by Thumbelina book and by langauge

thumbelina <- tidy_books_hca %>% 
  filter(book == "Thumbelina") %>% 
  filter(language == "English")

thumbelina
```

$~$

#### **Analysis `AFINN` and `LOUGHRAN`**

```{r}
afinn_hca <- thumbelina %>% 
  inner_join(get_sentiments("afinn")) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")
```



```{r}
loughran_hca <- thumbelina %>%
  inner_join(get_sentiments("loughran")) %>%
  filter(!is.na(sentiment)) %>%
  count(sentiment, sort = TRUE)
loughran_hca
```



```{r}
loughran_hca_word_counts <- thumbelina %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

loughran_hca_word_counts
```



```{r}
loughran_hca_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```



```{r}
hca_custom_stop_words <- bind_rows(tibble(word = c("the"),  
                                      lexicon = c("custom")), 
                               stop_words)

hca_custom_stop_words
```


$~$

#### **Wordclouds**

```{r}
tidy_books_hca %>%
  filter(book == "Thumbelina") %>% 
  filter(language == "English") %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```



```{r}
tidy_books_hca %>%
  filter(book == "Thumbelina") %>% 
  inner_join(get_sentiments("bing")) %>%
  filter(language == "English") %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

$~$

### **Conclusion:**

#### Both lexicons gave a variability for a sentiment analysis. For the code extension, I filtered the book "Thumbelina" which has more joyous words along with a couple of negatives. It was important to filter by language since the package contained 4 other languages. 

$~$

### **References:**
$~$

* Robinson, J. S. and D. (n.d.). 2 sentiment analysis with Tidy Data: Text mining with R. 2 Sentiment analysis with tidy data | Text Mining with R. Retrieved October 25, 2021, from https://www.tidytextmining.com/sentiment.html. 

* Get_sentiments: Get a tidy data frame of a single sentiment lexicon. RDocumentation. (n.d.). Retrieved October 28, 2021, from https://www.rdocumentation.org/packages/tidytext/versions/0.2.3/topics/get_sentiments.  

* EmilHvitfeldt. (n.d.). Emilhvitfeldt/hcandersenr: An R package for H.C. Andersens Fairy tales. GitHub. Retrieved October 29, 2021, from https://github.com/EmilHvitfeldt/hcandersenr. 





