---
title: "Data 607 Tidyverse CREATE Assignment"
author: "Leticia Salazar"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: pygments
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

[Kaggle.com](https://www.kaggle.com/johnsmith88/heart-disease-dataset)

* age
* sex
* chest pain type (4 values)
* resting blood pressure
* serum cholesterol in mg/dl
* fasting blood sugar > 120 mg/dl
* resting electrocardiographic results (values 0,1,2)
* maximum heart rate achieved
* exercise induced angina
* oldpeak = ST depression induced by exercise relative to rest
* the slope of the peak exercise ST segment
* number of major vessels (0-3) colored by flourosopy
* thal: 0 = normal; 1 = fixed defect; 2 = reversable defect


```{r}
#load libraries
library("tidyverse")
library("dplyr")
library("tidyr")
library("ggplot2")
```



```{r}
# load csv file

heart_disease <- read.csv("https://raw.githubusercontent.com/letisalba/Data-607/main/Week_9_HWK/Tidyverse/heart.csv", header = TRUE, na = ",")
head(heart_disease, n = 4)
```

```{r}
glimpse(heart_disease)
```

```{r}
summary(heart_disease)
```


```{r}
colSums(is.na(heart_disease))
```



