---
title: "Data 607 - Project 1"
author: "Leticia Salazar"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
collaborators: Chinedu Onyeka and Tyler Baker
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Install packages
#install.packages("tidyverse")
```


#### Project 1

#### In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .csv file (that could for example be imported into a SQL database) with the following information for all of the players:

* Player's Name
* Player's State
* Total Number of Points 
* Player's Pre-Rating
* Average Pre Chess Rating of Opponents

##### For the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605

##### 1605 was calculated by using the pre-tournament opponent's ratings of 1436, 1536, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played. 

##### **Objectives:** From the cross-tables, choose only the player's opponents and averge pre-rating of their opponents, both for players who played all of the scheduled games (8 points), and for players who had one or more unplayed games (e.g. byes, forfeits) (5 points). Are the average ratings presented to nearest full-point accuracy? (2 points)

##### Using the provided ELO calculation, determine each player's expected results (number of points), based on his or her pre-tournament rating, and the average pre-tournament rating for all of the player's opponents. Which player scored the most points relative to his or her expected results? (3 extra-credit points)

***

```{r}
#Load libraries
library(tidyverse)
library(stringr)
library(dplyr)
```

#### Loading Dataset
```{r}
#Load the dataset

theUrl <- "https://raw.githubusercontent.com/letisalba/Data-607-Project1/main/tournament.txt"

tournament <- read.delim(file = theUrl, header = FALSE, sep = "|", skip = 3) #Separating by "|" and skipping the first 3 rows

head(tournament)
```

```{r}
#Get a glimpse of the dataset
#Note that there's plenty of dashes and NA's which will be removed
glimpse(tournament)
```


#### Data Wrangling
```{r}
#Removing dashes
tournament2 <- tournament[!grepl("---", tournament[ ,1]), ]
head(tournament2)
```

#### Odd Rows Cleansing
```{r}
#For one player there is two rows that split between Player's name, points and rounds while the second row contains Player's state and pre/post ratings along with other information in regards to the rounds
#Start by splitting dataset into odd and even rows to be able to clean them individually

#Split dataset into odd row only

odd_rows <- tournament2[seq(1, nrow(tournament2), 2), ]

#Removing columns not needed in odd_rows
odd_rows <- odd_rows[, -c(1, 11)]

#Add column ID named Player Numbers
#This is going to help merge the two datasets later on
odd_rows$Player_Number <- seq(1, 64, length.out = dim(odd_rows)[1])



head(odd_rows)
```

```{r}
#Create new column names for even_rows dataset and rename
colnames(odd_rows) <- c("Player_Name", "Points", "Opp_1", "Opp_2", "Opp_3", "Opp_4", "Opp_5", "Opp_6", "Opp_7", "Player_Number")

#Reorder columns in 'odd_rows' to place Player_Number before Player_Name
odd_rows2 <- odd_rows[c("Player_Number", "Player_Name", "Points", "Opp_1", "Opp_2", "Opp_3", "Opp_4", "Opp_5", "Opp_6", "Opp_7")]
head(odd_rows2)
```

#### Even Rows Cleansing
```{r}
#Split dataset into even rows only
even_rows <- tournament2[seq(2, nrow(tournament2), 2), ]
#Remove columns not needed
even_rows <- even_rows[, -c(3:11)]

# Add column ID to match Player Numbers
# This is going to help merge the two datasets later on
even_rows$Player_Number <- seq(1, 64, length.out = dim(even_rows)[1])

head(even_rows)
```


```{r}
# Create new column names for even_rows dataset and rename
colnames(even_rows) <- c("State", "Ratings_Pre/Post", "Player_Number")

# Reorder columns in even_rows so that Player_Number is before State
even_rows2 <- even_rows[c("Player_Number", "State","Ratings_Pre/Post")]
head(even_rows2)
```


```{r}
#Join odd_rows and even_rows datasets by "Player_Number" to form one table
tournament_subset <- left_join(odd_rows, even_rows2, by = "Player_Number")

#Reorder columns 
tournament_subset2 <- tournament_subset[c("Player_Number", "Player_Name", "State", "Points", "Ratings_Pre/Post", "Opp_1", "Opp_2", "Opp_3", "Opp_4", "Opp_5", "Opp_6", "Opp_7")]
head(tournament_subset2)
```


#### Cleansing Ratings_Pre/Post column
```{r}
# Spliting Ratings_Pre/Post to get Pre and Post ratings alone
split_ratings <- str_split_fixed(even_rows2$`Ratings_Pre/Post`, "-", 2)
colnames(split_ratings) <- c("Pre-Rating", "Post-Rating")

#Remove numbers before "R:"
PreRating <- unlist(str_remove_all(split_ratings, "\\d+\\d [/]"))

#Extract Player's Pre-Ratings
PreRating <- unlist(str_extract_all(split_ratings, "R: \\d+\\d"))

#Remove "R:"
PreRating2 <- unlist(str_remove_all(PreRating, "R:"))

#Make values as.numeric
PreRatings <- as.numeric(PreRating2)
PreRatings
```


```{r}
#Define new values
New_PreRatings = c("1794", "1553", "1384", "1716"," 1655", "1686", "1649", "1641", "1411", "1365", "1712", "1663", "1666", "1610", "1220", "1604", "1629", "1600", "1564", "1595", "1563", "1555", "1363", "1229", "1745", "1579", "1552", "1507", "1602", "1522", "1494", "1441", "1449", "1399", "1438", "1355", "980", "1423", "1436", "1348", "1403", "1332", "1283", "1199", "1242", "377", "1362", "1382", "1291", "1056", "1011", "935", "1393", "1270", "1186", "1153", "1092",  "917",  "853",  "967", "955", "1530", "1175", "1163")

#cbind tables together with tournament_subset2
tournament_subset3 <-  cbind(tournament_subset2, New_PreRatings)

head(tournament_subset3)
```

```{r}
#Remove Ratings_Pre/Post Column
tournament_subset4 <- tournament_subset3[, -c(5)]

#Reorder column names so that New_PreRatings comes after Points and print new dataset
tournament_subset5 <- tournament_subset4[c("Player_Number", "Player_Name", "State", "Points", "New_PreRatings", "Opp_1", "Opp_2", "Opp_3", "Opp_4", "Opp_5", "Opp_6", "Opp_7")]
head(tournament_subset5)
```


#### Cleansing Opponents columns 1 - 7 
```{r}
#Start extracting Opponents columns into matrix to remove W, L, D from the Player Numbers
opponents <- tournament_subset5[6:12]
Opps <- matrix(str_extract(unlist(opponents), "\\d+"), ncol = 7)
head(Opps)
```


```{r}
#Name New Columns
opponents_col <- c("Opp1", "Opp2", "Opp3", "Opp4", "Opp5", "Opp6", "Opp7")

#Set as a dataframe and call in columns
Opps_df <- as.data.frame(Opps)
colnames(Opps_df) <- opponents_col

#Replace all NA's with 0
Opps_df[is.na(Opps_df)] <- 0

head(Opps_df)
```


```{r}
#cbind Opps_df together with tournament_subset5
tournament_subset6 <-  cbind(tournament_subset5, Opps_df)

#Remove columns 6 to 12 from tournament _subset6
tournament_subset7 <- tournament_subset6[, -c(6:12)]
head(tournament_subset7)
```


#### Computing Opponents Averages
```{r}
#Calculate Opponents Averages with a For Loop 
for(i in 1:nrow(tournament_subset7)){
  tournament_subset7$Opp_Avg_Ratings[i] <- round(mean(PreRatings[as.numeric(Opps_df[i,])], na.rm = TRUE))
}

head(tournament_subset7)
```

```{r}
#Remove columns 6 to 12 from tournament _subset7
chess_tournament <- tournament_subset7[, -c(6:12)]

#Print final dataset
head(chess_tournament, n = 6)
```

#### Write .CSV File
```{r}
#Write .csv file
write.csv(chess_tournament, file = "chess_tournament.csv")
```




