---
title: "Data 607 Assignment Week 9"
author: "Leticia Salazar"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: pygments
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment – Web APIs

##### The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis You’ll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

##### For this assignment I will be using [Top Stories API](https://developer.nytimes.com/docs/top-stories-product/1/overview) and focus on the section World.


***

### Libraries

```{r}
# install packages if you don't have them
#install.packages("rjson")


# load libraries
library("tidyverse")
library("jsonlite")
library("httr")
library("stringr")
library("ggplot2")
library("stringr")
library("dplyr")
```

### Connection to API

```{r}
# API Key
apikey <- "x3mOLM3nRnwjYVOuE89tloE62oC8NofC"

# Get the URL 
theURL <- paste("https://api.nytimes.com/svc/topstories/v2/world.json?api-key=", apikey)

topstories <- GET(theURL)

# Get status code
topstories$status_code
```

```{r}
summary(topstories)
```

```{r}
top_stories <- content(topstories, as = "text")
```


### Data Frame Conversion

```{r}
#Option 1
top_storiesdf <- top_stories %>% 
  fromJSON() %>% 
  as.data.frame()

#Get column names
colnames(top_storiesdf)
```

```{r}
#Option 2
top_storiesdf2 <- fromJSON(top_stories, flatten = TRUE)
top_storiesdf2 <- data.frame(top_storiesdf2$results, stringsAsFactors = FALSE)

#Get column names
colnames(top_storiesdf2)
```


```{r}
#Rename columns
colnames(top_storiesdf2) <- c("Section","Subsection", "Title", "Abstract", "URL", "URI", "Byline", "Item_Type", "Updated_Date", "Created_Date", "Published_Date", "Material_Type_Facet", "Kicker", "Des_Facet", "Org_Facet", "Per_Facet", "Geo_Facet", "Multimedia", "Short_Url")

#Drop columns not needed
top_storiesdf3 <- top_storiesdf2[, -c(12:18)]
```


### Data Analysis

```{r}
#Count of Section
Section <- top_storiesdf3%>%
  group_by(Section)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Section)
```


```{r}
#Count of Subsection
Subsection <- top_storiesdf3%>%
  group_by(Subsection)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Subsection)
```


### Plots

#### The bar plot below shows that plenty of the articles belong to the world section followed by briefing and science.

```{r}
# Bar Plot for Section
top_storiesdf3 %>% 
ggplot(aes(x = Section)) +
  geom_bar()
```


#### Within the next bar plot there is an unknown subsection that has the highest count followed by Asia and Europe.

```{r}
# Bar Plot for Subsection
top_storiesdf3 %>% 
ggplot(aes(x = Subsection)) +
  geom_bar()
```