---
title: "Data 607 Project2"
author: "Leticia Salazar"
collaborators: "Chinedu Onyeka, Tyler Baker"
date: "`r Sys.Date()`"
output:
  html_document: "default"
  pdf_document: "default"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### The goal of this assignment is to give you practive in preparing different datasets for downstram analysis work.

#### Your task is to:
(1) Choose any three of the "wide" datasets identified in the Week 6 Discussion items. (You may use your own dataset; please don't use my Sample Post dataset, since that was used in your Week 6 assignment!) **For each of the three chosen datasets**
* Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a "wide" structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
* Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
* Perform the analysis requested in the discussion item.
* Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.

(2) Please include in your homework submission, **for each of the three chosen datasets**:
* The URL to the .Rmd file in your GitHub repository, and 
* The URL for your rpubs.com web page.

_Note: My collaborators (Chinedu and Tyler) and I were each responsible to tidy one data set which we would then share and join into our project 2. Below you'll see the data set I chose to tidy._

***


#### Load the Libraries
```{r}
#Import libraries
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(hrbrthemes)
```

#### About the Dataset

##### The dataset was taken from Kaggle.com which was originally taken from an article called Hotel Booking Demands Datasets https://www.sciencedirect.com/science/article/pii/S2352340918315191. 

#### It contains 32 columns listed below with a brief description:

_Columns _                      | _Description_
--------------------------------|---------------------------------------------
Hotel                           | H1 = Resort Hotel, H2 = City Hotel
is_canceled                     | 1 = cancelled, 0 = not cancelled
lead_time                       | number of days that elapsed between entering the data into the Property Management System (PMS) and arrival date
arrival_date_year               | year of arrival date
arrival_date_month              | month of arrival date
arrival_date_week_number        | week number of year for arrival date
arrival_date_day_of_month       | day of arrival date
stays_in_weekend_nights         | number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel
stays_in_week_night             | number of weeknights (Monday - Friday) the guest stayed or booked to stay at the hotel
adults                          | number of adults
children                        | number of children
babies                          | number of babies
meal                            | types of meals booked: Undefined/SC = no meal package, BB = Bed & Breakfast, HB = Half board (breakfast and one other meal - usual dinner), FB = Full board (breakfast, lunch and dinner)
country                         | country of origin, categories are represented in the ISO 355-3.2013 format
market_segment                  | market segment designation, TA = Travel Agent, TO = Tour Operators
distribution_channel            | booking distribution channel, TA = Travel Agent, TO = Tour
is_repeated_guest               | Value indicating if the booking name was from a repeated guest (1) or not (0)
previous_cancellations          | number of previous bookings that were cancelled by the customer prior to the current booking
previous_booking_not_canceled   | number of previous bookings not cancelled by the customer prior to the current booking
reserved_room_type              | code of room type reserved; Code is presented instead of designation for anonymity reasons
assigned_room_type              | code for the type of room assigned to the booking. Sometimes the assigned room type differs from the reserved room type due to hotel operation reasons (e.g. overbooking) or by customer request. Code is presented instead of designation for anonymity reasons.
booking_changes                 | number of changes/amendments made to the booking from the moment the booking was entered on the PMS until the moment of check-in or cancellation
deposit_type                    | indication on if the customer made a deposit to guarantee the booking. No Deposit = no deposit was made, Non Refund = a deposit was made in the value of the total stay cost, Refundable = a deposit was made with a value under the total cost of stay
agent                           | ID of the travel agency that made the booking
company                         | ID of the company/entity that made the booking or responsible for paying the booking. ID is presented instead of designation for anonymity reasons
days_in_waiting_list            | number of days the booking was in the waiting list before it was confirmed to the customer
customer_type                   | type of booking; Contract = when the booking has an allotment or other type of contract associate to it, Group = when the booking is associated to a group, Transient = when the booking is not part of a group or contract, and is not associated to other transient booking, Transient-party = when the booking is transient, but is associated to at least other transient booking
adr                             | Average Daily Rate as defined by dividing the sum of all lodging transactions by the total number of staying nights
required_car_parking_spaces     | number of car parking spaces required by the customer
total_of_special_requests       | number of special requests made by the customer (e.g. twin bed or high floor)
reservation_status              | reservation last status; Canceled = booking was canceled by the customer, Check-Out = customer has checked in but already departed, No-Show = customer did not check-in and did inform the hotel of the reason why
reservation_status_date         | date at which the last status was set. This variable can be used in conjunction with the ReservationStatus to understand when was the booking canceled or when did the customer checked-out of the hotel



***
  
  
##### With this dataset I am looking to analyze the following:

* Cancellations of hotel with the arrival date month 
* Cancellations of hotel with arrival date year

##### I would also like to analyze the following

* What type of hotels do repeated guest book?
* What type of customers tend to be repeated guests?
* Do repeated guests have a preferred meal choice?
* Do guests have special requests?
* How do total stays compare with total costs at hotels?


***


#### Load the data set
```{r}
theUrl <- "https://raw.githubusercontent.com/letisalba/Data607_Project2/main/hotel_bookings%202.csv"

hotel_booking <- read.csv(file = theUrl, header = TRUE, sep = ",")
head(hotel_booking)
```


### Data Wrangling
```{r}
#Get column names
colnames(hotel_booking)
```


```{r}
#Rename column names for more clarity
colnames(hotel_booking) <-  c("Hotel", "Cancellation_Status", "Lead_Time", "Arrival_Year", "Arrival_Month", "Arrival_Week", "Arrival_Day_of_Month", "Stays_Weekend_Nights", "Stays_Week_Nights", "Adults", "Children", "Babies", "Meal", "Country", "Market_Segment", "Distribution_Channel", "Repeated_Guest", "Previous_Cancellations", "Prev_Non_Cancellations", "Reserved_Room", "Assigned_Room", "Booking_Changes", "Deposit_Type", "Agent", "Company", "Days_Waitlisted", "Customer_Type", "ADR", "Required_Parking_Spaces", "Special_Requests", "Reservation_Status", "Reservation_Status_Date")
head(hotel_booking)
```


```{r}
#Glimpse of data
glimpse(hotel_booking)
```


```{r}
#Summary of columns in data
summary(hotel_booking)
```


```{r}
#Convert columns to factor
hotel_bookings <- hotel_booking %>% 
   mutate(Cancellation_Status = as.factor(Cancellation_Status), 
          Meal = as.factor(Meal),
          Repeated_Guest = as.factor(Repeated_Guest),
         )
head(hotel_booking, n = 4)
```


```{r}
#Calculate total cost of stays and the total number of stays for each reservation and create two separate columns for each

hotel_bookings <- hotel_booking %>% 
  mutate(Stays_Total = Stays_Weekend_Nights + Stays_Week_Nights,
       Stays_Total_Cost = ADR * Stays_Total)
head(hotel_bookings, n = 4)
```


```{r}
#Find missing values
colSums(is.na(hotel_bookings))

#There are 4 values missing in children
```


```{r}
#Replace missing values in Children by using the Median
#By using the Median it will allow for those missing values to be added without trying to guess what number to input. Using the number 0 would have also worked without skewing the data.

children_median <- median(hotel_bookings$Children, na.rm = TRUE)

hotel_bookings$Children[is.na(hotel_bookings$Children)] <- children_median

head(hotel_bookings, n = 4)
```


```{r}
#Check to see if there's any other missing values
colSums(is.na(hotel_bookings))
```


```{r}
#Replace all NULL values in data set
hotel_bookings2 <- hotel_bookings %>% replace(.=="NULL", 0) # replace with 0
head(hotel_bookings2, n = 4)
```


```{r}
#Replacing values in Cancellation_Status column so that 0 = not cancelled and 1 = cancelled
hotel_bookings3<- recode(hotel_bookings2$Cancellation_Status,
                              "0" = "Not_Cancelled",
                              "1" = "Cancelled")    
head(hotel_bookings3)
```


```{r}
#Replacing values in Repeated_Guest column so that 0 = new guest and 1 = repeated guest
hotel_bookings4 <- recode(hotel_bookings2$Repeated_Guest,
                              "0" = "New_Guest",
                              "1" = "Repeated_Guest")    
head(hotel_bookings4)
```


```{r}
#cbind 3 tables 
hotelbookings <-  cbind(hotel_bookings2, hotel_bookings3, hotel_bookings4)

#Drop rows not needed
hotelbookings <- hotelbookings[, -c(2, 15:17, 20, 21, 23:25, 29, 32)]

head(hotelbookings)
```

```{r}
#Get column names
colnames(hotelbookings)
```


```{r}
#Rename column names
colnames(hotelbookings) <-  c("Hotel", "Lead_Time", "Arrival_Year", "Arrival_Month", "Arrival_Week", "Arrival_Day_of_Month", "Stays_Weekend_Nights", "Stays_Week_Nights", "Adults", "Children", "Babies", "Meal", "Country", "Previous_Cancellations", "Prev_Non_Cancellations", "Booking_Changes", "Days_Waitlisted", "Customer_Type", "ADR", "Special_Requests", "Reservation_Status", "Stays_Total", "Stays_Total_Cost", "Cancellation_Status", "Is_Repeated_Guest")

#Reorder columns
col_order <- c("Hotel", "Cancellation_Status", "Lead_Time", "Arrival_Year", "Arrival_Month", "Arrival_Week", "Arrival_Day_of_Month", "Stays_Weekend_Nights", "Stays_Week_Nights", "Stays_Total", "Stays_Total_Cost", "Adults", "Children", "Babies", "Meal", "Country", "Is_Repeated_Guest", "Previous_Cancellations", "Prev_Non_Cancellations", "Booking_Changes", "Days_Waitlisted", "Customer_Type", "ADR", "Special_Requests", "Reservation_Status")

hotelbookings <- hotelbookings[, col_order]
head(hotelbookings)
```


### Data Exploration


##### **City Hotels have the hightest number of bookings**
```{r}
#Count of Hotels
Hotel <- hotel_bookings2%>%
  group_by(Hotel)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Hotel)
```


##### **There has been more non cancellations than cancellations when booking**
```{r}
#Count of Cancellation_Status
Cancellation_Status <- hotel_bookings2%>%
  group_by(Cancellation_Status)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Cancellation_Status)
```


##### **August and July are the two highest months when hotels are booked**
```{r}
#Count of Arrival_Month
Arrival_Month <- hotel_bookings2%>%
  group_by(Arrival_Month)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Arrival_Month)
```


##### **Zero and two weekend nights are amongst the highest guest have stayed in a hotel**
```{r}
#Count of Stays_Weekend_Nights
Stays_Weekend <- hotel_bookings2%>%
  group_by(Stays_Weekend_Nights)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Stays_Weekend)
```


##### **Two and one week nights are the most guest stay or book to stay at the hotel.**
```{r}
#Count of Stays_Week_Nights
Stays_Week <- hotel_bookings2%>%
  group_by(Stays_Week_Nights)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Stays_Week)
```


##### **Bed and Breakfast is the most common meal option.**
```{r}
#Count of Meals
Meal <- hotel_bookings2%>%
  group_by(Meal)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Meal)
```


##### **Non repeated guest tend to book stays more often than repeated guests.**
```{r}
#Count of Repeated_Guest
Repeated_Guest <- hotel_bookings2%>%
  group_by(Repeated_Guest)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Repeated_Guest)
```


##### **Transient customers book the most hotel stays..**
```{r}
#Count on Customer_Type
Customer <- hotel_bookings2%>%
  group_by(Customer_Type)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Customer)
```


##### **Zero and 1 special requests are made by the customer**
```{r}
#Count on Special_Requests
Special_Requests <- hotel_bookings2%>%
  group_by(Special_Requests)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Special_Requests)
```


##### **When combining the stays for weekend nights and week nights, the two days is the highest number of stays.**
```{r}
#Count on Stays_Total
Stays_Total <- hotel_bookings2%>%
  group_by(Stays_Total)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Stays_Total)
```


##### **The most common total cost in a hotel stay is $124 (assuming the currency is dollars)**
```{r}
#Count on Stays_Total_Cost
Stays_Total_Cost <- hotel_bookings2%>%
  group_by(Stays_Total_Cost)%>%
  summarise(num=n())%>%
  arrange(desc(num))
head(Stays_Total_Cost)
```


### Data Analysis:


##### Cancellations of hotel with Arrival Month
_Cancellations in April is the highest._
```{r}
# Bar Plot 
hotel_bookings2 %>% 
ggplot(aes(x = Cancellation_Status, fill = Arrival_Month)) +
  geom_bar()
```


##### Cancellations of hotel with arrival date year
_The year 2016 had the most cancellations/non-cancellations._
```{r}
# Bar Plot 
hotel_bookings2 %>% 
ggplot(aes(x = Arrival_Year, fill = Cancellation_Status)) +
  geom_bar()
```


##### What type of hotels do repeated guest book?
_Repeated guests book City Hotels the most._
```{r}
# Bar Plot 
hotel_bookings2 %>% 
ggplot(aes(x = Repeated_Guest, fill = Hotel)) +
  geom_bar()
```


##### What type of customers tend to be repeated guests?
_Transient customers tend to be repeated guests._
```{r}
# Plot
hotel_bookings2%>%
  ggplot(aes(x = Repeated_Guest, fill = Customer_Type)) +
  geom_bar()
```


##### Do repeated guests have a preferred meal choice?
_Bed and Breakfast is the preferred meal of choice for repeated guest._
```{r}
# Plot
hotel_bookings2%>%
  ggplot(aes(x = Repeated_Guest, fill = Meal)) +
  geom_bar()
```


##### Do guests have special requests?
_Guests in general have no special requests during hotel stay._
```{r}
# Plot
hotel_bookings2%>%
  ggplot(aes(x = Special_Requests, fill = Repeated_Guest)) +
  geom_bar()
```


##### How do total stays compare with total costs at hotels?
_Whether it's a City or Resort hotel, majority of the booking stays are  between 0 to 20 days with a few outliers._
```{r}
#Scatter plot
hotel_bookings2 %>% 
ggplot(aes(x = Stays_Total, y = Stays_Total_Cost, color = Hotel)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm , color="red", se=FALSE) + # with linear trend
  theme_ipsum()
```


***
