---
title: "Data 607 Assignment Week 3"
author: "Leticia Salazar"
date: "`r Sys.Date()`"
output: "html_document"

---

##### Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to the problems below.  You may work in a small group, but please submit separately with names of all group participants in your submission.

#### 1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Install package
#install.packages("stringr")
```

```{r}
#Import libraries
library(tidyverse)
library(stringr)
library(dplyr)
```


```{r}
#Load the dataset

theUrl <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

majors_df <- read.csv(file = theUrl, header = TRUE, sep = ",")

head(majors_df)
```

```{r}
#Use str_detect to filter out majors containing "DATA" and "STATISTICS"
df <- majors_df
df %>%
  filter(str_detect(Major, "DATA|STATISTICS"))
```

#### 2. Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
#Transform from text string to character vector using str_c
fruits <- str_c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry", sep = ",")
fruits


# This is using str_extract_all

fruits <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry") 
fruits <- unlist(str_extract_all(fruits, pattern = "(.+?)+", simplify = FALSE))
fruits
```


The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#### 3. Describe, in words, what these expressions will match:

* (.)\1\1
* "(.)(.)\\2\\1"
* (..)\1
* "(.).\\1.\\1"
* "(.)(.)(.).*\\3\\2\\1"

```{r}
#Bringing up the fruits data frame to show the regex patterns
head(fruit)
```


* **_(.)\1\1: This first regex won't return anything._**

```{r}
#Assigning the pattern and using the fruit data frame to see what it returns with str_subset and str_match


#pattern <- (.)\1\1 
#fruit %>% 
  #str_subset(pattern)%>%
  #str_match(pattern)
```


* **_"(.)(.)\\2\\1":  This second regex returns fruits that have repeated pair of letters and returns the second letter first followed by the first letter._**

```{r}
#Assigning the pattern and using the fruit data frame to see what it returns with str_subset and str_match


#pattern <- "(.)(.)\\2\\1"
#fruit %>% 
  #str_subset(pattern)%>%
  #str_match(pattern)
```

* **_(..)\1:  This regex does not return anything._**

```{r}
#Assigning the pattern and using the fruit data frame to see what it returns with str_subset and str_match


#pattern <- (..)\1
#fruit %>% 
  #str_subset(pattern)%>%
  #str_match(pattern)
```

* **_"(.).\\1.\\1":  This regex returns the fruit that end and start with the same letter and returns that letter once._**


```{r}
#Assigning the pattern and using the fruit data frame to see what it returns with str_subset and str_match


#pattern <- "(.).\\1.\\1"
#fruit %>% 
  #str_subset(pattern)%>%
  #str_match(pattern)
```


* **_"(.)(.)(.).*\\3\\2\\1":  This last regex does not return anything._**

```{r}
#Assigning the pattern and using the fruit data frame to see what it returns with str_subset and str_match


#pattern <- "(.)(.)(.).*\\3\\2\\1"
#fruit %>% 
  #str_subset(pattern)%>%
  #str_match(pattern)
```


####  4. Construct regular expressions to match words that:

* Start and end with the same character.
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
str_view("anemia","(.)(.*)\\1$")
```

```{r}
str_view("church", "([A-Za-z][A-Za-z]).*\\1" )
```

```{r}
str_view("eleven", "([A-Za-z]).*\\1.*\\1.")
```



